---
title: "Foxe Basin Output"
output: html_notebook
---

## All Time Trends and Cluster Analsysis Using the Functions  #########################
## Again, But Better With Functions ###################################################

### Packages

```{r}
library(tidyverse)
library(mkac)
library(readxl)
```

Data


## Breakup ####################################################################################


```{r}
fname="FoxeBasin_Brkup.csv"
fname_coords="Foxe Basin Location Coordinates 3 proper.xlsx"
sea_ice_data=read_in(fname)
sea_ice_long=make_long(sea_ice_data)
```



## Time Trend Plots ###############################################

### Linear Trends ###############

```{r, fig.height=12, fig.width=12}
time_trend_plot_linear(sea_ice_long)
```
```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(method="lm", se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```



### Linear Trend Slopes - Magnitude of Trends ##########

```{r}
slopes=ls_slopes(sea_ice_long)
slopes
```

### Average slope of Linear Trends #########

```{r}
slopes %>% summarize(mean_slope=mean(linear_slope))
```

### Minimum and Maximum trends (slopes) observed by location, plus all of the positive slopes (for Breakup)


```{r}
slopes %>% 
  filter(linear_slope<0) %>% 
  filter(linear_slope==min(linear_slope) | linear_slope==max(linear_slope))
slopes %>% filter(linear_slope>0) # no rows means there are none of them
```




### Lowess Curves ###############

```{r, fig.height=12, fig.width=12}
time_trend_plot_lowess(sea_ice_long)
```

```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```





## Mann-Kendall and Theil-Sen #########################################################

### Mann-Kendall ##################################

```{r}
tab=mann_kendall_table(sea_ice_long)
tab
```

```{r}
tab %>% count(P_value_adj<=0.05)
```

### Theil-Sen ###############################################

```{r}
slopes=ts_slopes(sea_ice_long)
slopes
```

### Mean of Theil Sen Slope ########################

```{r}
slopes %>% summarize(mean=mean(theil_sen), sd=sd(theil_sen))
```




## Mann-Kendall Maps ###################################

```{r, fig.height=9, fig.width=9}
coordinates=read_coordinates(fname_coords)
mk_map(tab, coordinates, long_lim=c(72, 84), lat_lim=c(63, 71), longlegend=88, latlegend = 69)
```




## Theil-Sen Maps ###################################


```{r, fig.width=9, fig.height=9}
draw_ts_map(locations, xlim=c(-84, -72), ylim=c(63, 71), slopes$theil_sen, scaling=2.7)
```


## Cluster Analysis ###################################################################################

```{r}
sea_ice_data
```




### Ward's Method ##################################


##### Missings Included

```{r}
w_1=dendrogram(sea_ice_data, remove_missing = F, n_cluster=6) # with missing included
w_1$tree
w_1$clusters
```


##### Missings Excluded

```{r}
w_2=dendrogram(sea_ice_data, remove_missing = T, n_cluster=6) # with missing excluded
w_2$tree
w_2$clusters
```



```{r}
k_1=km(sea_ice_data, n_cluster=6)
k_1$scree
k_1$clusters
```




## Cluster Maps ##############################################

```{r}
library(rworldmap)
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
lons=c(-84,-72)
lats=c(63,71)
```


### Ward's Method (Missings Included)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=w_1$clusters)
```



### Ward's Method (Missings Excluded)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=w_2$clusters)
```




### K-means Method (Missings Excluded)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=k_1$clusters)
```




## Freeze-up ####################################################################################

```{r}
fname="FoxeBasin_Frzup.csv"
fname_coords="Foxe Basin Location Coordinates 3 proper.xlsx"
sea_ice_data=read_in(fname)
sea_ice_long=make_long(sea_ice_data)
```



## Time Trend Plots ###############################################

### Linear Trends ###############

```{r, fig.height=12, fig.width=12}
time_trend_plot_linear(sea_ice_long)
```
```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(method="lm", se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```



### Linear Trend Slopes - Magnitude of Trends ##########

```{r}
slopes=ls_slopes(sea_ice_long)
slopes
```

### Average slope of Linear Trends #########

```{r}
slopes %>% summarize(mean_slope=mean(linear_slope))
```

### Minimum and Maximum trends (slopes) observed by location, plus all of the negative slopes (for Freeze-up)


```{r}
slopes %>% 
  filter(linear_slope>0) %>% 
  filter(linear_slope==min(linear_slope) | linear_slope==max(linear_slope))
slopes %>% filter(linear_slope<0) # no rows means there are none of them
```






### Lowess Curves ###############

```{r, fig.height=12, fig.width=12}
time_trend_plot_lowess(sea_ice_long)
```

```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```





## Mann-Kendall and Theil-Sen #########################################################

### Mann-Kendall ##################################

```{r}
tab=mann_kendall_table(sea_ice_long)
tab
```

```{r}
tab %>% count(P_value_adj<=0.05)
```

### Theil-Sen ###############################################

```{r}
slopes=ts_slopes(sea_ice_long)
slopes
```

### Mean of Theil Sen Slope ########################

```{r}
slopes %>% summarize(mean=mean(theil_sen), sd=sd(theil_sen))
```




## Mann-Kendall Maps ###################################

```{r, fig.height=9, fig.width=9}
coordinates=read_coordinates(fname_coords)
mk_map(tab, coordinates, long_lim=c(72, 84), lat_lim=c(63, 71), longlegend=88, latlegend = 69)
```




## Theil-Sen Maps ###################################


```{r, fig.width=9, fig.height=9}
draw_ts_map(locations, xlim=c(-84, -72), ylim=c(63, 71), slopes$theil_sen, scaling=2.7)
```


## Cluster Analysis ###################################################################################

```{r}
sea_ice_data
```




### Ward's Method ##################################


##### Missings Included

```{r}
w_1=dendrogram(sea_ice_data, remove_missing = F, n_cluster=6) # with missing included
w_1$tree
w_1$clusters
```


##### Missings Excluded

```{r}
w_2=dendrogram(sea_ice_data, remove_missing = T, n_cluster=6) # with missing excluded
w_2$tree
w_2$clusters
```



```{r}
k_1=km(sea_ice_data, n_cluster=6)
k_1$scree
k_1$clusters
```




## Cluster Maps ##############################################

```{r}
library(rworldmap)
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
lons=c(-84,-72)
lats=c(63,71)
```


### Ward's Method (Missings Included)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=w_1$clusters)
```



### Ward's Method (Missings Excluded)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=w_2$clusters)
```




### K-means Method (Missings Excluded)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=k_1$clusters)
```


## Ice Free ####################################################################################

```{r}
fname="FoxeBasin_Icefree.csv"
fname_coords="Foxe Basin Location Coordinates 3 proper.xlsx"
sea_ice_data=read_in(fname)
sea_ice_long=make_long(sea_ice_data)
```



## Time Trend Plots ###############################################

### Linear Trends ###############

```{r, fig.height=12, fig.width=12}
time_trend_plot_linear(sea_ice_long)
```
```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(method="lm", se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```



### Linear Trend Slopes - Magnitude of Trends ##########

```{r}
slopes=ls_slopes(sea_ice_long)
slopes
```

### Average slope of Linear Trends #########

```{r}
slopes %>% summarize(mean_slope=mean(linear_slope))
```

### Minimum and Maximum trends (slopes) observed by location, plus all of the negative slopes (for Ice Free)


```{r}
slopes %>% 
  filter(linear_slope>0) %>% 
  filter(linear_slope==min(linear_slope) | linear_slope==max(linear_slope))
slopes %>% filter(linear_slope<0) # no rows means there are none of them
```






### Lowess Curves ###############

```{r, fig.height=12, fig.width=12}
time_trend_plot_lowess(sea_ice_long)
```

```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```





## Mann-Kendall and Theil-Sen #########################################################

### Mann-Kendall ##################################

```{r}
tab=mann_kendall_table(sea_ice_long)
tab
```

```{r}
tab %>% count(P_value_adj<=0.05)
```

### Theil-Sen ###############################################

```{r}
slopes=ts_slopes(sea_ice_long)
slopes
```

### Mean of Theil Sen Slope ########################

```{r}
slopes %>% summarize(mean=mean(theil_sen), sd=sd(theil_sen))
```




## Mann-Kendall Maps ###################################

```{r, fig.height=9, fig.width=9}
coordinates=read_coordinates(fname_coords)
mk_map(tab, coordinates, long_lim=c(72, 84), lat_lim=c(63, 71), longlegend=88, latlegend = 69)
```




## Theil-Sen Maps ###################################


```{r, fig.width=9, fig.height=9}
draw_ts_map(locations, xlim=c(-84, -72), ylim=c(63, 71), slopes$theil_sen, scaling=2.7)
```


## Cluster Analysis ###################################################################################

```{r}
sea_ice_data
```




### Ward's Method ##################################


##### Missings Included

```{r}
w_1=dendrogram(sea_ice_data, remove_missing = F, n_cluster=6) # with missing included
w_1$tree
w_1$clusters
```


##### Missings Excluded

```{r}
w_2=dendrogram(sea_ice_data, remove_missing = T, n_cluster=6) # with missing excluded
w_2$tree
w_2$clusters
```



```{r}
k_1=km(sea_ice_data, n_cluster=6)
k_1$scree
k_1$clusters
```




## Cluster Maps ##############################################

```{r}
library(rworldmap)
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
lons=c(-84,-72)
lats=c(63,71)
```


### Ward's Method (Missings Included)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=w_1$clusters)
```



### Ward's Method (Missings Excluded)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=w_2$clusters)
```




### K-means Method (Missings Excluded)


```{r, fig.height=9, fig.width=9}
draw_map(locations, xlim=lons, ylim=lats, clusters=k_1$clusters)
```