---
title: "Code for Cluster Analysis"
output: html_notebook
---

## packages

```{r}
library(tidyverse)
```


###############################################################################################################################################
###############################################################################################################################################
##################################################################################################################################################


## Breakup ##

## read in data

## file name

```{r}
fname="FoxeBasin_Brkup.csv"
```

## read in data

```{r}
sea_ice_data=read_csv(fname)
sea_ice_data %>% rename(year=X1) -> sea_ice_data
sea_ice_data
```

```{r}
n_cluster=6
```


inter-location distances

```{r}
sea_ice_data %>% select(-year) %>% t() %>% dist() -> d
d
```

change number of clusters to suit

```{r}
bu.hc=hclust(d,method="ward.D")
plot(bu.hc,xlab="missings included")
rect.hclust(bu.hc,n_cluster)
```

```{r}
cl.8.ward.with.missing=cutree(bu.hc,k=n_cluster)
cl.8.ward.with.missing
```

Now, remove any years with missing values

```{r}
sea_ice_data %>% drop_na()
```

```{r}
sea_ice_data %>% drop_na() %>%
  select(-year) %>% t() %>% dist() -> d2
```

ward again. Again change number of clusters to suit.

```{r}
bu.hc=hclust(d,method="ward.D")
plot(bu.hc,xlab="missings excluded")
rect.hclust(bu.hc,n_cluster)
```

again, for map

```{r}
cl.8.ward.without.missing=cutree(bu.hc,k=n_cluster)
cl.8.ward.without.missing
```

## K-means

Make scree plot

```{r}
sea_ice_data %>% drop_na() %>% select(-year) -> vv
tibble(clusters=2:15) %>% 
  mutate(km=map(clusters, ~kmeans(vv, ., nstart=40))) %>% 
  mutate(ss=map_dbl(km, "tot.withinss")) -> twss
ggplot(twss, aes(x=clusters, y=ss))+geom_point()+geom_line()
```

## insert your preferred value for number of clusters (eg 6)

```{r}
vv %>% t() %>% kmeans(n_cluster, nstart=40) -> kmm
kmclusters=enframe(kmm$cluster,name="location", value="cluster")
kmclusters
```



##################################################################################################################################################
###############################################################################################################################################

## Cluster Maps ##

```{r}
#install.packages("rworldmap")
library(rworldmap)
```

read in locations

```{r}
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
locations
```


```{r}
newmap=getMap(resolution="low")

#Foxe Basin Map with Locations

with(locations, -range(Longitude))
with(locations, range(Latitude))
lons=c(-84,-72)
lats=c(63,71)
with(locations,plot(newmap,xlim=lons,ylim=lats))

```

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=1,cex=1.7))
with(locations, text(-Longitude,Latitude,Location,pos=3))

```




##################################################################################################################################################

## Ward Clustering (With Missing) ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=cl.8.ward.with.missing,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```

##################################################################################################################################################

## Ward Clustering (Without Missing) ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=cl.8.ward.without.missing,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```


##################################################################################################################################################

## K-means Clustering (Without Missing)  ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=kmclusters$cluster,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```



##################################################################################################################################################
##################################################################################################################################################
##################################################################################################################################################

## Freeze-up ##

## read in data

## file name

```{r}
fname="FoxeBasin_Frzup.csv"
```

## read in data

```{r}
sea_ice_data=read_csv(fname)
sea_ice_data %>% rename(year=X1) -> sea_ice_data
sea_ice_data
```

```{r}
n_cluster=6
```


inter-location distances

```{r}
sea_ice_data %>% select(-year) %>% t() %>% dist() -> d
d
```

change number of clusters to suit

```{r}
bu.hc=hclust(d,method="ward.D")
plot(bu.hc,xlab="missings included")
rect.hclust(bu.hc,n_cluster)
```

```{r}
cl.8.ward.with.missing=cutree(bu.hc,k=n_cluster) # this 5 should agree with rect.hclust above
cl.8.ward.with.missing
```

Now, remove any years with missing values

```{r}
sea_ice_data %>% drop_na()
```

```{r}
sea_ice_data %>% drop_na() %>%
  select(-year) %>% t() %>% dist() -> d2
```

ward again. Again change number of clusters to suit.

```{r}
bu.hc=hclust(d,method="ward.D")
plot(bu.hc,xlab="missings excluded")
rect.hclust(bu.hc,n_cluster)
```

again, for map

```{r}
cl.8.ward.without.missing=cutree(bu.hc,k=n_cluster)
cl.8.ward.without.missing
```

## K-means

Make scree plot

```{r}
sea_ice_data %>% drop_na() %>% select(-year) -> vv
tibble(clusters=2:15) %>% 
  mutate(km=map(clusters, ~kmeans(vv, ., nstart=40))) %>% 
  mutate(ss=map_dbl(km, "tot.withinss")) -> twss
ggplot(twss, aes(x=clusters, y=ss))+geom_point()+geom_line()
```

## insert your preferred value for number of clusters (eg 6)

```{r}
vv %>% kmeans(n_cluster, nstart=40) -> kmm
kmclusters=enframe(kmm$cluster,name="location", value="cluster")
kmclusters
```



###############################################################################################################################################
###############################################################################################################################################

## Cluster Maps ##

```{r}
#install.packages("rworldmap")
library(rworldmap)
```

read in locations

```{r}
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
locations
```


```{r}
newmap=getMap(resolution="low")

#Foxe Basin Map with Locations

with(locations, -range(Longitude))
with(locations, range(Latitude))
lons=c(-84,-72)
lats=c(63,71)
with(locations,plot(newmap,xlim=lons,ylim=lats))

```

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=1,cex=1.7))
with(locations, text(-Longitude,Latitude,Location,pos=3))

```




##################################################################################################################################################

## Ward Clustering (With Missing) ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=cl.8.ward.with.missing,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```

##################################################################################################################################################

## Ward Clustering (Without Missing) ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=cl.8.ward.without.missing,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```


##################################################################################################################################################

## K-means Clustering (Without Missing)  ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=kmclusters$cluster,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```



##################################################################################################################################################
##################################################################################################################################################
##################################################################################################################################################


## Ice Free ##

## read in data

## file name

```{r}
fname="FoxeBasin_Icefree.csv"
```

## read in data

```{r}
sea_ice_data=read_csv(fname)
sea_ice_data %>% rename(year=X1) -> sea_ice_data
sea_ice_data
```

```{r}
n_cluster=6
```

inter-location distances

```{r}
sea_ice_data %>% select(-year) %>% t() %>% dist() -> d
d
```

change number of clusters to suit

```{r}
bu.hc=hclust(d,method="ward.D")
plot(bu.hc,xlab="missings included")
rect.hclust(bu.hc,n_cluster)
```

```{r}
cl.8.ward.with.missing=cutree(bu.hc,k=n_cluster)
cl.8.ward.with.missing
```

Now, remove any years with missing values

```{r}
sea_ice_data %>% drop_na()
```

```{r}
sea_ice_data %>% drop_na() %>%
  select(-year) %>% t() %>% dist() -> d2
```

ward again. Again change number of clusters to suit.

```{r}
bu.hc=hclust(d,method="ward.D")
plot(bu.hc,xlab="missings excluded")
rect.hclust(bu.hc,n_cluster)
```

again, for map

```{r}
cl.8.ward.without.missing=cutree(bu.hc,k=n_cluster)
cl.8.ward.without.missing
```

## K-means

Make scree plot

```{r}
sea_ice_data %>% drop_na() %>% select(-year) -> vv
tibble(clusters=2:15) %>% 
  mutate(km=map(clusters, ~kmeans(vv, ., nstart=40))) %>% 
  mutate(ss=map_dbl(km, "tot.withinss")) -> twss
ggplot(twss, aes(x=clusters, y=ss))+geom_point()+geom_line()
```

## insert your preferred value for number of clusters (eg 6)

```{r}
vv %>% kmeans(n_cluster, nstart=40) -> kmm
kmclusters=enframe(kmm$cluster,name="location", value="cluster")
kmclusters
```




###############################################################################################################################################

## Cluster Maps ##

```{r}
#install.packages("rworldmap")
library(rworldmap)
```

read in locations

```{r}
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
locations
```


```{r}
newmap=getMap(resolution="low")

#Foxe Basin Map with Locations

with(locations, -range(Longitude))
with(locations, range(Latitude))
lons=c(-84,-72)
lats=c(63,71)
with(locations,plot(newmap,xlim=lons,ylim=lats))

```

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=1,cex=1.7))
with(locations, text(-Longitude,Latitude,Location,pos=3))

```




##################################################################################################################################################

## Ward Clustering (With Missing) ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=cl.8.ward.with.missing,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```

##################################################################################################################################################

## Ward Clustering (Without Missing) ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=cl.8.ward.without.missing,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```


##################################################################################################################################################

## K-means Clustering (Without Missing)  ##

```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=kmclusters$cluster,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))

```




##################################################################################################################################################
##################################################################################################################################################

## Old Cluster Map Code ##

##################################################################################################################################################
##################################################################################################################################################



```{r}


make.map=function(my.map,v,colours=1:36)
{
  # translate given colours into appropriate length vector for 
  # plotting
  plot.cols=colours[v]
  plot(my.map,xlim=c(-90,-80),ylim=c(50,65),pch=19)
  pts=read.table("locations.txt",header=T)
  attach(pts)
  points=data.frame(number=1:36,long=-long,lat=lat)
  detach(pts)
  attach(points)
  points(long,lat,pch=19,col=plot.cols)
  text(long,lat,number,pos=4)  
  detach(points)
  w=levels(v)
  legend(-108,64,legend=w,col=colours,pch=19)
}



#install.packages("rworldmap")
library(rworldmap)

newmap=getMap(resolution="low")

#Hudson Strait Map with Locations

plot(newmap,xlim=c(-80,-63),ylim=c(57,66))
pts=read.table("locations2_HudsonStrait.txt",header=T)
head(pts)
attach(pts)
points=data.frame(number=Location,long=-Longitude,lat=Latitude)
detach(pts)
points
attach(points)
points(long,lat,pch=19,col=1)
text(long,lat,number,pos=3)

detach(points)


#Foxe Basin Map with Locations

plot(newmap,xlim=c(-84,-72),ylim=c(64,71))
pts=read.table("locations3_FoxeBasin.txt",header=T)
head(pts)
attach(pts)
points=data.frame(number=Location,long=-Longitude,lat=Latitude)
detach(pts)
points
attach(points)
points(long,lat,pch=19,col=1)
text(long,lat,number,pos=3)

detach(points)

##################################################################
#Jottings

# make a map with all points black and numbered

plot(newmap,xlim=c(-90,-80),ylim=c(50,65))
pts=read.table("locations.txt",header=T)
pts
attach(pts)
points=data.frame(number=1:36,long=-long,lat=lat)
points
detach(pts)
attach(points)
points(long,lat,pch=19)
text(long,lat,number,pos=4)
detach(points)

drawmap=function(map,points,clustering)
{
  title=deparse(substitute(clustering))
  plot(map,xlim=c(-90,-80),ylim=c(50,65))
  text(-100,55,title)
  attach(points)
  points(long,lat,pch=19,col=clustering,cex=2)
  text(long,lat,number,pos=4,cex=0.75)  
  detach(points)
  invisible()
}

drawmap(newmap,points,fake.cluster)

```


```

