---
Title: "Code for Lowess Curves, Mann-Kendall and Theil-Sen"
Output: html_notebook
---

## packages

```{r}
library(tidyverse)
library(mkac)
```



###################################################################################################################################################
###################################################################################################################################################

## Breakup ##

## file name

```{r}
fname="FoxeBasin_Brkup.csv"
```

## read in data

```{r}
sea_ice_data=read_csv(fname)
sea_ice_data %>% rename(year=X1) -> sea_ice_data
sea_ice_data
```

## time trends

```{r}
sea_ice_data %>% gather(location, day, -year) %>% 
  mutate(location=as.numeric(location)) -> sea_ice_long
sea_ice_long
```

time trend graphs

```{r fig.height=12, fig.width=12}
ggplot(sea_ice_long, aes(x=year, y=day))+geom_point()+
  geom_smooth(se=F) +
  facet_wrap(~location)
```

```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```


## Mann-Kendall and Theil-Sen

```{r}
sea_ice_long
```

```{r}
sea_ice_long %>% 
  drop_na() %>% 
  nest(-location) %>% 
  mutate(MK=map(data, ~as_tibble(kendall_Z_adjusted(.$day)))) %>% 
  unnest(MK) %>% 
  mutate(P_value_2=format(P_value, scientific=F)) %>% 
  mutate(P_value_adj_2=format(P_value_adj, scientific=F)) -> mann_kendall
mann_kendall
```

## counting significant ones

```{r}
mann_kendall %>% count(P_value_adj<=0.05)
```

```{r}
tss=function(d) {
  d %>% filter(!is.na(day)) -> dd
  theil_sen_slope(dd$day)
}

sea_ice_long %>% nest(-location) %>% 
  mutate(theil_sen=map_dbl(data, ~tss(.))) -> slopes
slopes
```
## mean of Theil Sen Slope

```{r}
slopes %>% summarize(mean=mean(theil_sen), sd=sd(theil_sen))
```


##############################################################################################################################################

## Mann-Kendall Significance Maps ##

read locations

```{r}
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
locations
```


```{r}
mann_kendall
```

```{r}
mann_kendall %>% 
  mutate(stars=cut(P_value_adj,breaks=c(0,0.01,0.05,0.10,1),labels=c("***","**","*","NS"))) %>% 
  mutate(plot_colour=case_when(
    P_value_adj<=0.01 ~ "blue",
    P_value_adj<=0.05 ~ "red",
    P_value_adj<=0.10 ~ "yellow",
    TRUE ~ "green"
  )) -> mann_kendall2
```

```{r}
mann_kendall2
mann_kendall2 %>% distinct(plot_colour, stars) %>% 
  mutate(rank=case_when(
    plot_colour == "blue" ~ 1,
    plot_colour == "red" ~ 2,
    plot_colour == "yellow" ~ 3,
    plot_colour == "green" ~ 4
  )) %>% 
  arrange(rank) -> legend_map
legend_map
```


```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=mann_kendall2$plot_colour,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))
legend(-88,69,legend=legend_map$stars,col=legend_map$plot_colour,pch=19, cex=2.3)
```





###################################################################################################################################################
###################################################################################################################################################

## Freeze-up ##

## file name

```{r}
fname="FoxeBasin_Frzup.csv"
```

## read in data

```{r}
sea_ice_data=read_csv(fname)
sea_ice_data %>% rename(year=X1) -> sea_ice_data
sea_ice_data
```

## time trends

```{r}
sea_ice_data %>% gather(location, day, -year) %>% 
  mutate(location=as.numeric(location)) -> sea_ice_long
sea_ice_long
```

time trend graphs

```{r fig.height=12, fig.width=12}
ggplot(sea_ice_long, aes(x=year, y=day))+geom_point()+
  geom_smooth(se=F) +
  facet_wrap(~location)
```

```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```


## Mann-Kendall and Theil-Sen

```{r}
sea_ice_long
```

```{r}
sea_ice_long %>% 
  drop_na() %>% 
  nest(-location) %>% 
  mutate(MK=map(data, ~as_tibble(kendall_Z_adjusted(.$day)))) %>% 
  unnest(MK) %>% 
  mutate(P_value_2=format(P_value, scientific=F)) %>% 
  mutate(P_value_adj_2=format(P_value_adj, scientific=F)) -> mann_kendall
mann_kendall
```

## counting significant ones

```{r}
mann_kendall %>% count(P_value_adj<=0.05)
```

```{r}
tss=function(d) {
  d %>% filter(!is.na(day)) -> dd
  theil_sen_slope(dd$day)
}

sea_ice_long %>% nest(-location) %>% 
  mutate(theil_sen=map_dbl(data, ~tss(.))) -> slopes
slopes
```
## mean of Theil Sen Slope

```{r}
slopes %>% summarize(mean=mean(theil_sen), sd=sd(theil_sen))
```



##############################################################################################################################################

## Mann-Kendall Significance Maps ##

read locations

```{r}
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
locations
```


```{r}
mann_kendall
```

```{r}
mann_kendall %>% 
  mutate(stars=cut(P_value_adj,breaks=c(0,0.01,0.05,0.10,1),labels=c("***","**","*","NS"))) %>% 
  mutate(plot_colour=case_when(
    P_value_adj<=0.01 ~ "blue",
    P_value_adj<=0.05 ~ "red",
    P_value_adj<=0.10 ~ "yellow",
    TRUE ~ "green"
  )) -> mann_kendall2
```

```{r}
mann_kendall2
mann_kendall2 %>% distinct(plot_colour, stars) %>% 
  mutate(rank=case_when(
    plot_colour == "blue" ~ 1,
    plot_colour == "red" ~ 2,
    plot_colour == "yellow" ~ 3,
    plot_colour == "green" ~ 4
  )) %>% 
  arrange(rank) -> legend_map
legend_map
```


```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=mann_kendall2$plot_colour,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))
legend(-88,69,legend=legend_map$stars,col=legend_map$plot_colour,pch=19, cex=2.3)
```





###################################################################################################################################################
###################################################################################################################################################


## Ice Free ##

## file name

```{r}
fname="FoxeBasin_Icefree.csv"
```

## read in data

```{r}
sea_ice_data=read_csv(fname)
sea_ice_data %>% rename(year=X1) -> sea_ice_data
sea_ice_data
```

## time trends

```{r}
sea_ice_data %>% gather(location, day, -year) %>% 
  mutate(location=as.numeric(location)) -> sea_ice_long
sea_ice_long
```

time trend graphs

```{r fig.height=12, fig.width=12}
ggplot(sea_ice_long, aes(x=year, y=day))+geom_point()+
  geom_smooth(se=F) +
  facet_wrap(~location)
```

```{r}
sea_ice_long %>% 
  nest(-location) %>% 
  mutate(g=map2(data, location, ~ggplot(.x, aes(x=year, y=day))+geom_point()+geom_smooth(se=F)+ ggtitle(as.character(.y)))) %>% 
  pull(g)
```


## Mann-Kendall and Theil-Sen

```{r}
sea_ice_long
```

```{r}
sea_ice_long %>% 
  drop_na() %>% 
  nest(-location) %>% 
  mutate(MK=map(data, ~as_tibble(kendall_Z_adjusted(.$day)))) %>% 
  unnest(MK) %>% 
  mutate(P_value_2=format(P_value, scientific=F)) %>% 
  mutate(P_value_adj_2=format(P_value_adj, scientific=F)) -> mann_kendall
mann_kendall
```

## counting significant ones

```{r}
mann_kendall %>% count(P_value_adj<=0.05)
```

```{r}
tss=function(d) {
  d %>% filter(!is.na(day)) -> dd
  theil_sen_slope(dd$day)
}

sea_ice_long %>% nest(-location) %>% 
  mutate(theil_sen=map_dbl(data, ~tss(.))) -> slopes
slopes
```
## mean of Theil Sen Slope

```{r}
slopes %>% summarize(mean=mean(theil_sen), sd=sd(theil_sen))
```



##############################################################################################################################################

## Mann-Kendall Significance Maps ##

read locations

```{r}
library("readxl")
locations=read_excel("Foxe Basin Location Coordinates 3 proper.xlsx", skip=1)
locations
```


```{r}
mann_kendall
```

```{r}
mann_kendall %>% 
  mutate(stars=cut(P_value_adj,breaks=c(0,0.01,0.05,0.10,1),labels=c("***","**","*","NS"))) %>% 
  mutate(plot_colour=case_when(
    P_value_adj<=0.01 ~ "blue",
    P_value_adj<=0.05 ~ "red",
    P_value_adj<=0.10 ~ "yellow",
    TRUE ~ "green"
  )) -> mann_kendall2
```

```{r}
mann_kendall2
mann_kendall2 %>% distinct(plot_colour, stars) %>% 
  mutate(rank=case_when(
    plot_colour == "blue" ~ 1,
    plot_colour == "red" ~ 2,
    plot_colour == "yellow" ~ 3,
    plot_colour == "green" ~ 4
  )) %>% 
  arrange(rank) -> legend_map
legend_map
```


```{r fig.height=9, fig.width=9}
with(locations,plot(newmap,xlim=lons,ylim=lats))
with(locations, points(-Longitude,Latitude,pch=19,col=mann_kendall2$plot_colour,cex=2.3))
with(locations, text(-Longitude,Latitude,Location,pos=3,cex=1.7))
legend(-88,69,legend=legend_map$stars,col=legend_map$plot_colour,pch=19, cex=2.3)
```



